---
title: "Day 1"
format: html
editor: visual
---

[AdventOfCode](https://jrosell.github.io/AdventOfCode/) \> 2022

## Part 1: How many Calories are being carried by the Elf carrying the most Calories?

I manually downloaded my personal [day 1 input file](https://adventofcode.com/2022/day/1) as a logged user, and here I get the data with blank lines as NA and "calories" as column name:

```{r}
library(tidyverse)
data <- read_csv(
        here::here("2022/01_input"),
        col_names = c("calories"),
        show_col_types = FALSE,
        skip_empty_rows = FALSE
    )
data %>% print(n = 30)
```

I identify elfs for each row, starting by 1:

```{r}
data %>% 
    mutate(elf_id = 1 + cumsum(is.na(calories)))

```

I sum the calories carried by each elf:

```{r}
data %>% 
    mutate(elf_id = 1 + cumsum(is.na(calories))) %>% 
    group_by(elf_id) %>%
    summarise(calories = sum(calories, na.rm = TRUE))

```

I sort the data to find the Elf carrying the most Calories and select it.

```{r}
data %>% 
    mutate(elf_id = 1 + cumsum(is.na(calories))) %>% 
    group_by(elf_id) %>%
    summarise(calories = sum(calories, na.rm = TRUE)) %>% 
    arrange(-calories) %>%
    head(1)

```

## Part 2: How many calories are carried by the top three Elves carrying the most Calories?

```{r}
data %>% 
    mutate(elf_id = 1 + cumsum(is.na(calories))) %>% 
    group_by(elf_id) %>%
    summarise(calories = sum(calories, na.rm = TRUE)) %>% 
    arrange(-calories) %>% 
    head(3) %>% 
    summarize(top3 = sum(calories))
```
