---
title: "Day 2"
format: html
editor: visual
---

[AdventOfCode](https://jrosell.github.io/AdventOfCode/) \> 2022

## Part 1: What would your total score be if everything goes exactly according to your strategy guide?

I manually downloaded my personal [day 2 input file](https://adventofcode.com/2022/day/2) as a logged user, and here I get the data.

```{r}
library(tidyverse)
data <- read_delim(
        here::here("2022/02_input"),
        col_names = c("opponent", "me"),
        show_col_types = FALSE,
        delim=" "
    )
data %>% print(n = 30)
```

I define the rules to play and I codify the data according to my entities representation:

```{r}
entities <- c(
    A = "rock", B = "paper", C = "scissors",
    X = "rock", Y = "paper", Z = "scissors"
)
data %>% 
    mutate(opponent = entities[opponent]) %>% 
    mutate(me = entities[me])
```

I add a column for each turn result according to the game rules:

```{r}
lost <- 0
draw <- 3
win <- 6
rules <- tribble(
    ~opponent,  ~me,         ~my_outcome,
    "rock",     "scissors",  lost,                  
    "scissors", "paper",     lost,
    "paper",    "rock",      lost,
    "rock",     "rock",      draw,
    "scissors", "scissors",  draw,
    "paper",    "paper",     draw,
    "scissors", "rock",      win,
    "paper",    "scissors",  win,
    "rock",     "paper",     win,
)
data %>% 
    mutate(opponent = entities[opponent]) %>% 
    mutate(me = entities[me])
```

I get the score for each turn:

```{r}
lose <- 0
draw <- 3
win <- 6

picked_score <- c(
  "rock" = 1,
  "paper" = 2, 
  "scissors" = 3
)
rules <- tribble(
    ~opponent,  ~me,         ~my_outcome,
    "rock",     "scissors",  lose,                  
    "scissors", "paper",     lose,
    "paper",    "rock",      lose,
    "rock",     "rock",      draw,
    "scissors", "scissors",  draw,
    "paper",    "paper",     draw,
    "scissors", "rock",      win,
    "paper",    "scissors",  win,
    "rock",     "paper",     win,
)
turns <- data %>% 
    mutate(opponent = entities[opponent]) %>% 
    mutate(me = entities[me]) %>% 
    mutate(picked_score = picked_score[me]) %>% 
    left_join(rules, by = c("opponent", "me")) %>% 
    mutate(score = picked_score + my_outcome) 

turns
```

And the total score:

```{r}
sum(turns$score)
```

**Part 2: What would your total score be if everything goes exactly according to your strategy guide?**

I codify entites to get my turn score.

```{r}
entities <- c(
    A = "rock", B = "paper", C = "scissors",
    X =  lose, Y = draw, Z = win
)
data %>% 
    mutate(opponent = entities[opponent]) %>% 
    mutate(me = entities[me])
```

I get my strategy table.

```{r}
picked_score <- c(
  "rock" = 1,
  "paper" = 2, 
  "scissors" = 3
)
strategy <- rules %>% 
    mutate(picked_score = picked_score[me]) %>% 
    mutate(score = my_outcome + picked_score) %>% 
    transmute(opponent, me = my_outcome, score)

strategy
```

```{r}
turns <- data %>% 
    mutate(opponent = entities[opponent]) %>% 
    mutate(me = as.integer(entities[me])) %>% 
    left_join(strategy, by = c("opponent", "me"))

turns
```

And the total score:

```{r}
sum(turns$score)
```
