---
title: "Day 3"
format: html
editor: visual
---

[AdventOfCode](https://jrosell.github.io/AdventOfCode/) \> 2022

## Part 1: Find the item type that appears in both compartments of each rucksack. What is the sum of the priorities of those item types?

I manually downloaded my personal [day 3 input file](https://adventofcode.com/2022/day/3) as a logged user, and here I get the data.

```{r}
library(tidyverse)
data <- 
    read_csv(
        here::here("2022/03_input"),
        col_names = c("rucksack"),
        show_col_types = FALSE
    )
data %>% print(n = 30)
```

I tokenize rucksack characters and split them in half:

```{r}
data %>% 
    rowwise() %>% 
    mutate(rucksack = str_split(rucksack, "")) %>% 
    mutate(half = length(rucksack)/2) %>% 
    mutate(first = list(head(rucksack, half))) %>% 
    mutate(second = list(tail(rucksack, half)))
```

Find the item type that appears in both compartments of each rucksack

```{r}
data %>% 
    rowwise() %>% 
    mutate(rucksack = str_split(rucksack, "")) %>% 
    mutate(half = length(rucksack)/2) %>% 
    mutate(first = list(head(rucksack, half))) %>% 
    mutate(second = list(tail(rucksack, half))) %>% 
    mutate(problem = intersect(first, second))
```

I find the priority for each rucksack:

```{r}
item_types <- c(letters, LETTERS)

data %>% 
    rowwise() %>% 
    mutate(rucksack = str_split(rucksack, "")) %>% 
    mutate(half = length(rucksack)/2) %>% 
    mutate(first = list(head(rucksack, half))) %>% 
    mutate(second = list(tail(rucksack, half))) %>% 
    mutate(problem = intersect(first, second)) %>% 
    ungroup() %>% 
    mutate(priority = match(problem, item_types))

```

Here the full code:

```{r}
library(tidyverse)

item_types <- c(letters, LETTERS)

common_half <- function(x) {
  half <- length(x) / 2
  intersect(
    head(x, half), tail(x, half)
  )
}

read_csv(
        here::here("2022/03_input"),
        col_names = c("rucksack"),
        show_col_types = FALSE
    ) %>% 
    mutate(rucksack = str_split(rucksack, "")) %>% 
    rowwise() %>% 
    mutate(problem = common_half(rucksack)) %>% 
    ungroup() %>% 
    mutate(priority = match(problem, item_types)) %>% 
    summarise(result = sum(priority))
```

## Part 2: Find the item type that corresponds to the badges of each three-Elf group. What is the sum of the priorities of those item types?

I make groups for each 3 lines:

```{r}
data %>% 
  mutate(rucksack = str_split(rucksack, "")) %>% 
  group_by(elf_group = (row_number() + 2) %/% 3)
```

I find the item type that corresponds to the badges of each three-Elf group:

```{r}
data %>% 
  mutate(rucksack = str_split(rucksack, "")) %>% 
  group_by(elf_group = (row_number() + 2) %/% 3)  %>% 
  summarise(problem = reduce(rucksack, intersect))

```

Here the result:

```{r}
data %>% 
    mutate(rucksack = str_split(rucksack, "")) %>% 
    group_by(elf_group = (row_number() + 2) %/% 3)  %>% 
    summarise(problem = reduce(rucksack, intersect)) %>% 
    mutate(priority = match(problem, item_types)) %>% 
    summarise(result = sum(priority))
```

Here the full code:

```{r}
library(tidyverse)

item_types <- c(letters, LETTERS)

read_csv(
        here::here("2022/03_input"),
        col_names = c("rucksack"),
        show_col_types = FALSE
    ) %>% 
    mutate(rucksack = str_split(rucksack, "")) %>% 
    group_by(elf_group = (row_number() + 2) %/% 3)  %>% 
    summarise(problem = reduce(rucksack, intersect)) %>% 
    mutate(priority = match(problem, item_types)) %>% 
    summarise(result = sum(priority))
```
