---
title: "Day 6"
format: html
editor: visual
---

[AdventOfCode](https://jrosell.github.io/AdventOfCode/) \> 2022

## Part 1: How many characters need to be processed before the first start-of-packet marker is detected?

I manually downloaded my personal [day 6 input file](https://adventofcode.com/2022/day/6) as a logged user, and here I get the data in a more appropriate shape.

```{r}
library(tidyverse)

data <- 
    read_csv(
        here::here("2022/06_input"),
        col_names = c("data"),
        show_col_types = FALSE
    ) %>% 
    pull()

head(data)
```

I build a helper function and check some examples:

-   `bvwbjplbgvbhsrlpgdmjqwftvncz`: first marker after character *`5`*

-   `nppdvjthqldpwncqszvftbrmjlhg`: first marker after character *`6`*

-   `nznrnfrfntjfmvfwmzdfjlvtqnbhcprsg`: first marker after character *`10`*

-   `zcfzfwzzqfrljwzlrfnpqdbhtmscgvjw`: first marker after character *`11`*

```{r}
library(slider)

start_distinct_marker <- function(data, n) {
    data %>% 
        str_split_1("") %>%
        slide(~.x, .before = n - 1) %>% 
        map(function(x){
            n_distinct(x) == n
        }) %>% 
        which.max()
}

"bvwbjplbgvbhsrlpgdmjqwftvncz" %>%
    start_distinct_marker(4)

"nppdvjthqldpwncqszvftbrmjlhg" %>%
    start_distinct_marker(4)

"nznrnfrfntjfmvfwmzdfjlvtqnbhcprsg" %>%
    start_distinct_marker(4)

"zcfzfwzzqfrljwzlrfnpqdbhtmscgvjw" %>%
    start_distinct_marker(4)
```

Here the result:

```{r}
data %>% start_distinct_marker(4)
```

## Part 2: How many characters need to be processed before the first start-of-message marker is detected?

Here the full code for both parts:

```{r}
library(tidyverse)
library(slider)

data <- 
    read_csv(
        here::here("2022/06_input"),
        col_names = c("data"),
        show_col_types = FALSE
    ) %>% 
    pull() 

start_distinct_marker <- function(data, n) {
    data %>% 
        str_split_1("") %>% 
        slide(~.x, .before = n - 1) %>% 
        map(function(x){
            n_distinct(x) == n
        }) %>% 
        which.max()
}

data %>% start_distinct_marker(4)
data %>% start_distinct_marker(14)
```
